---
title: "Corvallis Climate"
author: "Medelin Kant"
date: "10/11/2021"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
library(ggplot2)
library(dplyr)
library(plotly)
library(hrbrthemes)
library(patchwork)
library(dplR)
library(stats)
library(utils)
```
Converting characters to dates
```{r, results='hide'}
#read data file
MDClimate <- read.csv("ClimateData_wDates.csv")
head(MDClimate)
dates <- MDClimate$Date
head(dates)
str(dates)
#convert dates from character class to dates class
ConvertedDate <- as.Date(dates)
#check to see if it worked
class(ConvertedDate)
```

Monthly Precipitation 1953-2021
```{r}
str(MDClimate)
PRCP <- MDClimate$PRCP

p <- MDClimate %>%
ggplot( aes(x=ConvertedDate, y=PRCP)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ylab("precipitation (in)") +
  xlab("January 1953 - July 2021") +
    theme_ipsum()

p <- ggplotly(p)
p
```

Monthly Average Temperature
```{r}
str(MDClimate)

TAVG <- MDClimate$TAVG

q <- MDClimate %>%
ggplot( aes(x=ConvertedDate, y=TAVG)) +
    geom_line(color="indianred4") +
    ylab("Average Temperature (F)") +
  xlab("January 1953 - July 2021") +
    theme_ipsum()

q <- ggplotly(q)
q
```


Monthly Minimum Temperature 
```{r}
str(MDClimate)

TMIN <- MDClimate$TMIN

q <- MDClimate %>%
ggplot( aes(x=ConvertedDate, y=TMIN)) +
    geom_line(color="midnightblue") +
    ylab("Minimum Temperature (F)") +
  xlab("January 1953 - July 2021") +
    theme_ipsum()

q <- ggplotly(q)
q


```

Monthly Maximum Temperatute 
```{r}
TMAX <- MDClimate$TMAX

x <- MDClimate %>%
ggplot( aes(x=ConvertedDate, y=TMAX)) +
    geom_line(color="sienna1") +
    ylab("Maximum Temperature (F)") +
  xlab("January 1953 - July 2021") +
    theme_ipsum()

x <- ggplotly(x)
x
```
# Time series with climate data and RWI 

p1 <- ggplot(MDClimate, aes(x=ConvertedDate, y=PRCP)) +
  geom_line(color="#69b3a2", size=2) +
  ggtitle("Precipitation") +
  theme_ipsum()
  
p1

PSMEcores <- read.rwl('MCores.rwl')
PSMEcores.rwi <- detrend(rwl = PSMEcores, method = "ModNegExp")

PSMEcores.rwi

Site1.rwi <- PSMEcores.rwi[c(1:68),c(1, 4:5, 7, 9, 12, 16:17, 19:20)]
Site2.rwi <- PSMEcores.rwi[c(1:68),c(2:3,6, 8, 10:11, 13:15,18)]
Live.rwi <- PSMEcores.rwi[c(1:68), c(1:3, 5, 7:8, 11:13, 16)]
Dead.rwi <- PSMEcores.rwi[c(1:68), c(4, 6, 9:10, 14:15, 17:20)]

PSMEcores.crn <- chron(PSMEcores.rwi, prefix = "PSM")
PSMEcores.crn

Site1.crn <- chron(Site1.rwi, prefix ="S1")
Site2.crn <- chron(Site2.rwi, prefix ="S2")
Live.crn <- chron(Live.rwi, prefix = "L")
Dead.crn <- chron(Dead.rwi, prefix = "D")

Years <- PSMEcores.crn[c(1:68), c(1)]
PSMstd <- PSMEcores.crn[c(1:68), c(2)]

ggplot(PSMEcores.crn, aes(x=Years, y=PSMstd)) +
  geom_line(color="grey",size=2) +
  ggtitle("RWI") +
  theme_ipsum()
  
p2

