---
title: "Tree Rings"
author: "Medelin Kant"
date: "10/12/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Libraries
```{r}
library(dplR)
library(stats)
library(utils)
library(ggplot2)
```

Reading Tree Ring Data
```{r}
# read data
PSMEcores <- read.rwl('MCores.rwl')
# how many years and series there are
dim(PSMEcores)
# the series IDs
colnames(PSMEcores)
# first few years
head(rownames(PSMEcores))
# the class should be "rwl" and "data.frame"
class(PSMEcores)

```

Describing and Plotting Ring-Width Data
```{r}
rwl.report(PSMEcores)
Site1 <- PSMEcores[c(1:68),c(1, 4:5, 7, 9, 12, 16:17, 19:20)]
Site2 <- PSMEcores[c(1:68),c(2:3,6, 8, 10:11, 13:15,18)]
Years <- PSMEcores[c(1:68), c(1)]
PSMEcores
#spaghetti plot
plot(PSMEcores, plot.type="spag")
spag.plot(Site1, zfac = 1, useRaster = FALSE, res = 150)
spag.plot(Site2, zfac = 1, useRaster = FALSE, res = 150)
plot(Site1, plot.type="spag")
plot(Site2, plot.type="spag")

```


Detrending Ring Widths to RWI
```{r}
PSMEcores.rwi <- detrend(rwl = PSMEcores, method = "ModNegExp")
# check to see that both data.frames have same dimensions after standardization 
dim(PSMEcores)
dim(PSMEcores.rwi)
head(PSMEcores.rwi)
PSMEcores.rwi

# "ar1" = first-order autocorrelation 
rwi.stats(PSMEcores.rwi)

Site1.rwi <- PSMEcores.rwi[c(1:68),c(1, 4:5, 7, 9, 12, 16:17, 19:20)]
Site2.rwi <- PSMEcores.rwi[c(1:68),c(2:3,6, 8, 10:11, 13:15,18)]
Live.rwi <- PSMEcores.rwi[c(1:68), c(1:3, 5, 7:8, 11:13, 16)]
Dead.rwi <- PSMEcores.rwi[c(1:68), c(4, 6, 9:10, 14:15, 17:20)]

S1Live.rwi <- PSMEcores.rwi[c(1:68), c(1, 5, 7, 12, 16)]
S1Dead.rwi <- PSMEcores.rwi[c(1:68), c(4, 9, 17, 19:20)]

S2Live.rwi <- PSMEcores.rwi[c(1:68), c(2:3, 8, 11, 13)]
S2Dead.rwi <- PSMEcores.rwi[c(1:68), c(6,10, 14:15, 18)]
```


Descriptive Statistics 
```{r}
#raw ring width stats
rwl.stats(PSMEcores)
# RWI stats
rwl.stats(PSMEcores.rwi)

# Interseries function gives the measure of average interseries correlation that is different fron the rbar measurements from rwi.stats. Correlations are calculated serially between each tree-ring series and a master chronology built from all the other series in the rwl object. 
PSMEcores.rho <- interseries.cor(PSMEcores.rwi, prewhiten=TRUE, method="spearman")

# for all 20 series
PSMEcores.rho[1:20, ]

#mean for all series
MeanRWI <- mean(PSMEcores.rho[, 1])


PSMEcores.crn <- chron(PSMEcores.rwi, prefix = "PSM")
Site1.crn <- chron(Site1.rwi, prefix ="S1")
Site2.crn <- chron(Site2.rwi, prefix ="S2")
Live.crn <- chron(Live.rwi, prefix = "L")
Dead.crn <- chron(Dead.rwi, prefix = "D")
S1Live.crn <- chron(S1Live.rwi, prefix = "S1L")
S1Dead.crn <- chron(S1Dead.rwi, prefix = "S1D")
S2Live.crn <- chron(S2Live.rwi, prefix = "S2L")
S2Dead.crn <- chron(S2Dead.rwi, prefix = 'S2D')

View(PSMEcores.crn)


plot(PSMEcores.crn, xlab='Year', ylab='RWI')
plot(Site1.crn, xlab='Year', ylab='RWI')
plot(Site2.crn, xlab='Year', ylab='RWI')
plot(Live.crn, xlab='Year', ylab='RWI')
plot(Dead.crn, xlab='Year', ylab='RWI')

plot(S1Live.crn, xlab='Year', ylab='RWI', ylim=c(0.4, 1.7))
plot(S1Dead.crn, xlab='Year', ylab='RWI', ylim=c(0.4, 1.7))
plot(S2Live.crn, xlab='Year', ylab='RWI'
plot(S2Dead.crn, xlab='Year', ylab='RWI')

```

Building a Mean Value Chronology 
```{r} 
# after detrending, build a chronology by averaging across the years of the rwi object. "chron" uses Tukey's biweight robust mean (average that is unaffected by outliers)
PSMEcores.crn <- chron(PSMEcores.rwi, prefix= "CAM")

# This object has the same number of rows as rwi object. dimensions are chronology first and sample depth (number of series abailbale in that year) second. 
dim(PSMEcores.rwi)
dim(PSMEcores.crn) 

#chronology plot with smoothing spline added
plot(PSMEcores.crn, add.spline=TRUE, nyrs=20)
```


BAI 

BAI.in <- bai.in(rwl=PSMEcores)
summary(BAI.in)

BAI.out <- bai.out(rwl = PSMEcores)
summary(BAI.out)

data(MCores.rwl)

data(PSMEcores)

data(MCores.d2pith)
foo <- bai.in(rwl = gp.rwl, d2pith = gp.d2pith)
foo.crn <- chron(foo)
yrs <- time(foo.crn)
plot(yrs, foo.crn[, 1], type = "n",
     xlab = "Year", ylab = expression(mm^2))
lines(yrs, foo.crn[, 1], col = "grey", lty = "dashed")
lines(yrs, ffcsaps(foo.crn[, 1], nyrs = 32), col = "red", lwd = 2)

